Interactive Filter and Display of Hillary Clinton's Emails: \\A Cautionary Tale of Metadata

### 2017-03-28 VAST_submission/vgtc_journal_latex/paper.tex
Changed Chris-specific TODOs

\subsection{Data extraction and cleaning}
The raw HTML of each message was programmatically downloaded from the Wikileaks archive using \texttt{R} packages  \texttt{RCurl}~\cite{RCurl2016package} and \texttt{XML}~\cite{XML2016package}.  This took several hours and required \TODO{Chris: mentioned etiquette concerns here?  Timeout?}
Once downloaded, the HTML data were processed using the \texttt{R} packages \texttt{tm}~\cite{tm2008paper, tm2017package}, \texttt{stringr}~\cite{stringr2016package}, and \texttt{SnowballC}~\cite{snowballc2014package}.

@@

\subsection{Metadata}
For each email, from the HTML we extract as best we can the identity of the persons sending and receiving the email as well as the date and time at which the email was processed by the server. \TODO{Chris:  Can you mention which fields we did or did not actually use? i.e. To: From" CC: BCC: etc.?}

The HTML was constructed from email printed out, redacted, and then provided as pdfs by the State Department.  Consequently, detailed email header information as would normally be available electronically is mostly missing.   All time stamps appear to be the local date and time at which the server sent or received that email (e.g. no time zone or other source time or IP chain information is available).  Moreover, because of redaction (typically FOIA exemption B6 ~\cite{FOIA}),  sender and receiver emails may have truncated domains,  contain only the person's name, or be missing altogether.  In cases where both \texttt{From} and \texttt{To} are entirely missing, but there is an email chain within the message, we  \TODO{Chris?  What happens?}  (e.g. see \texttt{https://wikileaks.org/clinton-emails/emailid/31599}).

@@

Where possible, the email addresses have also been parsed so that they may be classified into one of four categories: those which are \texttt{.gov},  those which are \texttt{.mil},  those which are identifiable as coming from a domain that is neither  \texttt{.gov} nor  \texttt{.mil}, and those whose domain was not identifiable from the data.
\subsection{Redaction information}
Emails that are redacted are marked as such by the string ``RELEASE IN PART'' and by the presence of one or more of nine FOIA exemption codes B1, B2, \ldots, B9  marking the place where text is missing (redacted).  This provides two further pieces of quasi-metadata (since it is not actual email content) that can be used in analysis.
\subsection{Content information}
The entire (redacted) content of each email is available on the State Department and Wikileaks sites and the user is encouraged to view it there (the pdf forms are more informative, especially in appreciating the extent of the redactions).  To provide some coarse statistical summaries of the content, all word tokens are extracted and partially processed to reduce their number.  For example,  ``stopwords'', as identified by the\texttt{stopwords} function from  were removed.   \TODO{Chris: I see from EmailFunctions.R that you only use the default stopwords() and an additional list.  Maybe you should use union(stopwords(), stopwords(kind="SMART"), yourList) ?}  The remaining words were stemmed by the \texttt{stemDocument} function from \texttt{tm} and \texttt{SnowballC}.   While neither the stopword lists nor the stemming tools are particularly well tuned to this corpus of emails, they nevertheless provide some insight as to the topics covered.

>>>>

\subsection{Data extraction and cleaning}
The raw HTML of each message was programmatically downloaded from the Wikileaks archive using \texttt{R} packages  \texttt{RCurl}~\cite{RCurl2016package} and \texttt{XML}~\cite{XML2016package}.  This took several hours and required the use of manually inserted system pauses to prevent timeouts in the connection, most likely due to Wikileaks DDoS protection software. Besides avoiding such timeouts, these pauses are considered web-scraping etiquette and best practice.
Once downloaded, the HTML data were processed using the \texttt{R} packages \texttt{tm}~\cite{tm2008paper, tm2017package}, \texttt{stringr}~\cite{stringr2016package}, and \texttt{SnowballC}~\cite{snowballc2014package}.

@@

\subsection{Metadata}
For each email, from the HTML we extract as best we can the identity of the persons sending and receiving the email as well as the date and time at which the email was processed by the server. The fields used were the address to, address from, contact name to, contact name from, subject line, and time. As well, forwarding chains of email addresses were captured through the identification of any to or from fields followed by emails within the text. In cases where no contact name was present the address was substituted. When the address was missing no imputation was completed. Carbon copy information was not extracted.

The HTML was constructed from email printed out, redacted, and then provided as pdfs by the State Department.  Consequently, detailed email header information as would normally be available electronically is mostly missing.   All time stamps appear to be the local date and time at which the server sent or received that email (e.g. no time zone or other source time or IP chain information is available).  Moreover, because of redaction (typically FOIA exemption B6 ~\cite{FOIA}),  sender and receiver emails may have truncated domains,  contain only the person's name, or be missing altogether.  In cases where both \texttt{From} and \texttt{To} are entirely missing, but there is an email chain within the message, we do not impute values. \TODO{I could implement the earlier mentioned forward chain extraction, I thought I had already implemented it but it seems I did not}  (e.g. see \texttt{https://wikileaks.org/clinton-emails/emailid/31599}).

@@

Where possible, the email addresses have also been parsed so that they may be classified into one of four categories: those which are \texttt{.gov},  those which are \texttt{.mil},  those which are identifiable as coming from a domain that is neither  \texttt{.gov} nor  \texttt{.mil}, and those whose domain was not identifiable from the data.
\subsection{Redaction information}
Emails that are redacted are marked as such by the string ``RELEASE IN PART'' and by the presence of one or more of nine FOIA exemption codes B1, B2, \ldots, B9  marking the place where text is missing (redacted).  This provides two further pieces of quasi-metadata (since it is not actual email content) that can be used in analysis.
\subsection{Content information}
The entire (redacted) content of each email is available on the State Department and Wikileaks sites and the user is encouraged to view it there (the pdf forms are more informative, especially in appreciating the extent of the redactions).  To provide some coarse statistical summaries of the content, all word tokens are extracted and partially processed to reduce their number.  For example,  ``stopwords'', as identified by the\texttt{stopwords} function under either the "en" and "SMART" settings, or a set of custom stopwords, were removed. The remaining words were stemmed by the \texttt{stemDocument} function from \texttt{tm} and \texttt{SnowballC}.   While neither the stopword lists nor the stemming tools are particularly well tuned to this corpus of emails, they nevertheless provide some insight as to the topics covered.
